---
title: "Individual Project"
author: "Adyasha Mishra"
date: "2024-03-02"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

HOW DOES QUALITY OF SLEEP GET AFFECTED? Checking through various factors such as Physical Activity Level, Age, Gender, Daily Steps. 
```{r}
library(readr)
Sleep_Health_Lifestyle <- read_csv("~/Downloads/Business Intelligence/Sleep_Health_Project/Sleep Health Lifestyle.csv")

```



```{r}
library(ggplot2)
```


```{r}

# Create a ggplot scatterplot comparing Quality of Sleep and Physical Activity Level
ggplot(Sleep_Health_Lifestyle, aes(x = `Quality of Sleep`, y = `Physical Activity Level`, color = factor(Gender))) +
  geom_point() + # Add points to the plot
  geom_smooth(method = "lm", se = FALSE) + # Add linear trend line without standard error band
  labs(x = "Quality of Sleep", y = "Physical Activity Level", color = "Gender") + # Add axis and legend labels
  ggtitle("Comparison of Quality of Sleep and Physical Activity by Gender") + # Add title
  theme_minimal() # Set minimal theme

```
For males, as the physical acitvity level increases, quality of sleep increases. We see a positive correlation here. 
For females, as the physical activity decreases, quality of sleep increases. 


```{r}
library(dplyr)
avg_quality_sleep <- Sleep_Health_Lifestyle %>%
  group_by(Age) %>%
  summarize(avg_quality_sleep = mean(`Quality of Sleep`, na.rm = TRUE))

# Plot the average quality of sleep against age with a trend line
ggplot(avg_quality_sleep, aes(x = Age, y = avg_quality_sleep)) +
  geom_point(color = "#8da0cb", size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "red") + # Add trend line
  labs(x = "Age", y = "Average Quality of Sleep") +
  ggtitle("Average Quality of Sleep Across Age Groups") +
  theme_minimal()


```
We can interpret that average quality of sleep increases as Age increases


```{r}
ggplot(Sleep_Health_Lifestyle, aes(x = `Daily Steps`, y = `Quality of Sleep`, size = `Quality of Sleep`, color = `Physical Activity Level`)) +
  geom_point(alpha = 0.7) + # Add points to the plot with transparency
  scale_size_continuous(range = c(3, 10)) + # Set the range of bubble sizes
  labs(x = "Daily Steps", y = "Quality of Sleep", color = "Physical Activity Level", size = "Quality of Sleep") + # Add axis labels
  ggtitle("Relationship between Daily Steps, Quality of Sleep, and Physical Activity Level") + # Add title
  theme_minimal() + # Set minimal theme
  scale_color_gradient(low = "blue", high = "red") # Set color gradient
```
As Physical Activity increases with daily steps, we see that the quality of sleep is better and increases significantly. 



```{r}
# Load the required libraries
library(rpart)
library(rpart.plot)

# Check the structure of your dataset
str(Sleep_Health_Lifestyle)

# Create and train the Decision Tree model
tree_model <- rpart(`Quality of Sleep` ~ `Physical Activity Level`, data = Sleep_Health_Lifestyle)
print(tree_model)

# Summary of the model
summary(tree_model)

# Plot the decision tree
rpart.plot(tree_model, type = 4, extra = 1)



# Make predictions on the training data
predictions <- predict(tree_model, Sleep_Health_Lifestyle)

# Compare predictions with actual values
comparison <- data.frame(Actual = Sleep_Health_Lifestyle$`Quality of Sleep`, Predicted = predictions)
head(comparison)




```
The decision tree suggests that higher levels of physical activity are generally associated with better quality of sleep.

If "Physical Activity Level" < 52.5:
Further splits based on different thresholds:
< 31: Predicted quality of sleep ≈ 7.35.
31 - 41: Predicted quality of sleep ≈ 6.52.
≥ 41: Predicted quality of sleep ≈ 5.25.
If "Physical Activity Level" ≥ 52.5:
Further splits:
< 87.5: Predicted quality of sleep ≈ 7.95.
≥ 87.5: Predicted quality of sleep ≈ 7.07.

